MS Technical Paper: \\ Placement Algorithms for Heterogeneous FPGAs

### 2025-08-19 latex/paper/sections/12-future_work.tex
need to review AP_2012

As noted in Section~\ref{sec:placement}, replacing SA with AP requires splitting the placement stage into two substages:  
(1) \emph{global placement}, which determines continuous target positions using analytical optimization, and  
(2) \emph{detailed placement} (legalization), which snaps objects to valid \texttt{Site}s while adhering to hardware constraints.  
Global placement can be performed using any number of off-the-shelf (OTS) analytical solvers.

@@

The star model transforms the hyperedge into a set of 2-pin edges connecting the incident sink pins to the source pin.
The clique net model transforms the hyperdedge into a local fully connected network of 2-pin edges.
We could preserve the directionality of the edges for possible future optimizations, but since HPWL minimization is not directly affected by the directionality of nets, we can just reduce the directed hypergraph into an undirected graph.
Figure \ref{fig:net_model} shows an example of a hyperedge reduction using the star model and clique model.

@@

Here, module placements in the \(x\) and \(y\) directions are captured by the placement vectors \( \boldsymbol{x} = (x_1, x_2, ..., x_n) \) and \( \boldsymbol{y} = (y_1, y_2, ..., y_n) \).
Each 2-pin edge has a weight \( w_{i,j} \) and is calculated depending on the net model used.

@@

\begin{equation}
    \boldsymbol{\Phi} (\boldsymbol{x}) = \frac{1}{2} \boldsymbol{x^T} \boldsymbol{Q_x} \boldsymbol{x} + \boldsymbol{c_x^T} \boldsymbol{x} + const.
    \label{equ:quadratic}
\end{equation}

% use these formulations
% HPWL_AP_2000
% HPWL_AP_2012

\subsection{Variations on Packing}
Our current placer follows a \textbf{Site-centric} approach that resembles that of the old Xilinx ISE - the predecessor to Vivado used in the 2000s.
These days, Vivado performs \textbf{BEL-centric} placement without necessarily locking \texttt{Cell}s into \texttt{Site}s, allowing for a higher granularity of movement of \texttt{Cell}s. 
Enabling BEL-centric moveme
nt as opposed to Site-centric movement can improve HPWL minimization, but will add much more complexity to the packing and placement process to ensure robustness with respect to hardware constraints.

@@

\subsection{Add Hard Macro Support}

>>>>

As noted in Section~\ref{sec:placement}, replacing SA with AP requires splitting the placement stage into two substages:  
\emph{global placement}, which determines continuous target positions using analytical optimization, and  
\emph{detailed placement} (legalization), which snaps objects to valid \texttt{site}s while adhering to hardware constraints.  
Global placement can be performed using any number of off-the-shelf (OTS) analytical solvers.

@@

The star model transforms the hyperedge into a set of 2-pin edges connecting the incident sink pins to the source pin.
The clique net model transforms the hyperdedge into a local fully connected network of 2-pin edges between the vertices incident to the original hyperedge.
This can lead to higher quality results but requires a quadratic growth in edges.
We could preserve the directionality of the edges for possible future optimizations, but since HPWL minimization is not directly affected by the directionality of nets, we can just reduce the directed hypergraph into an undirected graph.
Figure \ref{fig:net_model} shows an example of a hyperedge reduction using the star model and clique model.
The weight of each edge corresponds inversely to the wirelength of the net it represents.

@@

Here, module placements in the \(x\) and \(y\) directions are captured by the placement vectors \( \boldsymbol{x} = (x_1, x_2, ..., x_n) \) and \( \boldsymbol{y} = (y_1, y_2, ..., y_n) \).
The weight can be calculated depending on the net model used.
Placers using the star net model typically assign the weight as the inverse of the HPWL of the net, as shown in (\ref{equ:weight}).
Placers using the clique model or the Bound2Bound (B2B) model include a \(p-1\) term in the denominator as shown in (\ref{equ:weight_linearized}), where \(p\) is the number of pins on the original hyperedge, to de-emphasize high-fanout nets like the clock and reset nets.

@@

\begin{equation}
    \boldsymbol{\Phi} (\boldsymbol{x}) = \frac{1}{2} \boldsymbol{x}^T \boldsymbol{Q_x} \boldsymbol{x} + \boldsymbol{c_x}^T \boldsymbol{x} + const.
    \label{equ:quadratic}
\end{equation}

\subsection{Variations on Packing}
Our current placer follows a \textbf{Site-centric} approach that resembles that of the old Xilinx ISE - the predecessor to Vivado used in the 2000s.
These days, Vivado performs \textbf{BEL-centric} placement without necessarily locking \texttt{Cell}s into \texttt{Site}s, allowing for a higher granularity of movement of \texttt{Cell}s. 
Enabling BEL-centric moveme
nt as opposed to Site-centric movement can improve HPWL minimization, but will add much more complexity to the packing and placement process to ensure robustness with respect to hardware constraints, particularly with intra-\texttt{Site} routing.

@@

\subsection{Add Hard Macro Support}
fdsa

\subsection{Parallelization}
fdsa
